Layered Formal Verification of a TCP Stack

### 2021-06-04 SecDev21/paper.tex
Clement's feedback

The original CycloneTCP library provides no guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, has to be taken into account. The reason for this is two-fold. First, although TCP user functions can trigger state transitions, the actual transitions are done by other library functions called from TCP user functions. Second, other parts of the library can trigger state transitions during and between TCP user function calls, which can affect the intermediate and final states that a TCP user function can exhibit. The TCP user functions were fully translated to SPARK. SPARK bindings were also provided for most of the rest of the library's C functions to allow their invocation from the SPARK code.

>>>>

While being extensively tested against industry standard regression test suites, the original CycloneTCP library provides no formal guarantees that the implementation respects the TCP functional specification. Thus, there is no assurance that the implementation will only allow valid transitions between the different states that a TCP session can exhibit. We aim to use the SPARK technology to verify that the implementation respects the TCP automaton, given in \Cref{Fig:statemachine}. The work is focused on hardening the API of the user task, mainly the high-level user functions, to verify the state transitions. To achieve this, the rest of the TCP protocol, not related to the user task, has to be taken into account. The reason for this is two-fold. First, although TCP user functions can trigger state transitions, the actual transitions are done by other library functions called from TCP user functions. Second, other parts of the library can trigger state transitions during and between TCP user function calls, which can affect the intermediate and final states that a TCP user function can exhibit. The TCP user functions were fully translated to SPARK. SPARK bindings were also provided for most of the rest of the library's C functions to allow their invocation from the SPARK code.
